Asymmetric continuous prediction bands for assessing the validity of smooth biomechanical curve data

### 2021-09-12 paper.Rmd
Edit uncertainty plots

### TODO
- Alternativtitel: Comparing different methods for constructing continuous prediction bands
- Exemplarische Paper für die jeweiligen simulierten Datensätze raussuchen
- Weitere synthetische Daten siehe Paper Robinson et al. (2021)
- Evtl. simulierte Datensätze um Beispiele aus https://mjskay.github.io/ggdist/articles/lineribbon.html erweitern (insbesondere das erste Beispiel mit x-versetzten Gausskurven)!
- Simulierte Daten durch Funktionen ersetzen? Eigentlich aktuell eher unnötig, da die Zeitreihen ja eigentlich bereits mithilfe von Funktionen modelliert wurden. Zudem wird nicht interpoliert, abgeleitet o. Ä. ... letztlich für die aktuelle Idee des Papers eher ungeeignet (plus man umgeht Probleme wie mögliche Glättungsartefakte, oder die Definition eines functional mean).
- Falls doch FDA, dann evtl. als eigenständiges Skript forken
- Begründung warum Prädiktionsintervalle anstelle von Konfidenzintervallen
<!-- - Umbennung entire curve in "All-or-nothing" (entlehnt von Juul et al.) -->
- .bib file anlegen
- Begriff spatiotemporal überdenken (wie ist das im Kontext FDA definiert)
- Ratkowsky besorgen
- Begründung warum Prädiktionsintervalle und nicht Konfidenzintervalle (theoretisch könnte man beides machen; Prädiktionsintervalle scheinen mir aber eher das zu sein, was man will)

### TODO read (again)
- https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x
- Robinson, Vanrenterghem, Pataky (2021) Sample size estimation for biomechanical waveforms: Current practice, recommendations and a comparison to discrete power analysis
- Olshen et al. (1989) Gait analysis and the bootstrap
- Curvewise point and interval summaries for tidy data frames of draws from distributions https://mjskay.github.io/ggdist/reference/curve_interval.html
- Horvath, Kokoszka & Reeder (2013) Estimation of the mean of functional time series and a two-sample problem
- Generell die Kooperationspaper von Hörmann und Kokoszka zu FDA Statistik,insbesondere 'Inference for Functional Data with Applications' (2012)
- Ratkowsky 1990 Handbook of nonlinear regression models

@@

source("example_data.R")
# source("fdaDelta.R")
source("pick_subwise_curves.R")
source("draw_clusters.R")
source("floa_rcb.R")
source("floa_point.R")
source("floa_roislien.R")
source("plot_loa.R")
source("get_coverage.R")
source("get_coverage_fraction.R")
source("get_coverage_singlecurve.R")
source("get_coverage_singlecurve_fraction.R")
source("crossval_coverage.R")
source("crossval_coverage_fraction.R")
source("singlecurve_coverage.R")
source("singlecurve_coverage_fraction.R")
source("plot_cov_ver.R")
```

@@

* Randomized Cluster Bootstrap (FLoA~RCB~)
  + n = length(subjects) random strides from all strides (FLoA~RCB_v1~)
  + One random stride per subject (FLoA~RCB_v2~)
  + Fetch a SINGLE random stride from all strides (FLoA~RCB_v3~)

* Pointwise Gaussian intervals (FLoA~Point~)
  + FLoA~Point~ are calculated using mean and standard deviations (derived from linear mixed effects models)

* Roislien-like intervals (FLoA~Roslien~)
  + (Get one random stride from each subject ONCE and boot-
       strap the resulting sample (of length (n=length(subjects))

@@

#### Smooth curves with (missing) peaks

```{r echo = FALSE, warning = FALSE, message = FALSE}
data <- example_data(dat = "shock", dir.data)

@@

Leave-one out method to estimate the achieved coverage

Prediction intervals are calculated from size.dataset = number.subjects - 1 and then validated against the left out subjects curves for five different methods:

* FLoA~RCB_v1~   : n = length(subjects) random strides from all strides

@@

* FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
       strap the resulting sample (of length (n=length(subjects))

@@

<!-- $$ -->
<!-- RMSE_{delta} = \sqrt{\frac{1}{N}(target_1 - pedal_1)^2 + (target_2 - pedal_2)^2 + ... + (target_N - pedal_N)^2} -->
<!-- $$ -->

<!-- $$ -->
<!-- RMSE_{pedal} = \sqrt{\frac{1}{n}[(pedal_1 - \mu)^2 + (m_2 - \mu)^2 + ... + (pedal_N - \mu)^2]}, where \mu = \frac{1}{N}(pedal_1 + ... + pedal_N) -->
<!-- $$ -->

<!-- $$ -->
<!-- Z = \{z_1, z_2, z_3,...,z_i\} -->
<!-- $$ -->

@@

data <- example_data(dat = "smooth", dir.data)
floa <- floa_rcb(data, n.boot, ver = "v2")
plot_loa(data, floa, central.tendency = "mean")

data <- example_data(dat = "smooth_trend", dir.data)
floa <- floa_rcb(data, n.boot, ver = "v2")
plot_loa(data, floa, central.tendency = "mean")

data <- example_data(dat = "non_gaussian", dir.data)
floa <- floa_rcb(data, n.boot, ver = "v2")
plot_loa(data, floa, central.tendency = "mean")

data <- example_data(dat = "shock", dir.data)
floa <- floa_rcb(data, n.boot, ver = "v2")
plot_loa(data, floa, central.tendency = "mean")

data <- example_data(dat = "shift", dir.data)
floa <- floa_rcb(data, n.boot, ver = "v2")
plot_loa(data, floa, central.tendency = "mean")
```

@@

* 1. FLoA~RCB_v1~ : n = length(subjects) random strides from all strides

* 2. FLoA~RCB_v2~ : One random stride per subject

* 3. FLoA~RCB_v3~ : Fetch a SINGLE random stride from all strides

* 4. FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
       strap the resulting sample (of length (n=length(subjects))

* 5. FLoA~Point~ : Pointwise B & A Limits of Agreement (SD derived from linear mixed effects model)

@@

* b: Mean percentage of curve points within the respective FLoA: The fraction of points along a curve within the limits is calculated and averaged across all curves of a subject.

@@

a. (coverage of the entire curve)

```{r warning = FALSE, message = FALSE, eval = FALSE}
get_coverage_singlecurve <- function (curve, floa) {
  # Curve is a numerical vector
  # floa is a "matrix" "array" containing "lower" and "upper" limits

  # Calculate coverage (entire curves within the percentile boundaries)
  # ----------------------------------------------------------------------------
  lwr.bnd <- floa["lower", ]
  upr.bnd <- floa["upper", ]

  outside <- 0

  # Compare curves with upper and lower boundaries
  below.thresh <- curve < lwr.bnd
  above.thresh <- curve > upr.bnd

  points.outside <- sum(above.thresh) + sum(below.thresh)

  coverage <- ifelse(points.outside > 0, 0, 1)

  return(coverage)
}
```

@@

covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
covered.curves.percent <- data.frame(t(covered.curves.percent))
colnames(covered.curves.percent) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")

@@

b. (mean percentage of covered curve points)

```{r warning = FALSE, message = FALSE, eval = FALSE}
get_coverage_singlecurve_fraction <- function (curve, floa) {
  # Curve is a numerical vector
  # floa is a "matrix" "array" containing "lower" and "upper" limits

  # Calculate coverage (curve points within the percentile boundaries)
  # ----------------------------------------------------------------------------
  lwr.bnd <- floa["lower", ]
  upr.bnd <- floa["upper", ]

  outside <- 0

  # Compare difference curves with upper and lower boundaries
  below.thresh <- curve$value < lwr.bnd
  above.thresh <- curve$value > upr.bnd

  points.outside <- sum(above.thresh) + sum(below.thresh)

  # Divide by 101 to adjust for 101 data points
  coverage <- 1 - (points.outside / 101)
  coverage <- round(coverage, digits = 2)

  return(coverage)
}
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
# Leave-one curve out approach
# ------------------------------------------------------------------------------
cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
plot_cov_ver(cover.cross.fraction.singlecurve)

@@

a. (coverage of the entire curve)

@@

covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
covered.curves.percent <- data.frame(t(covered.curves.percent))
colnames(covered.curves.percent) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")

b. (mean percentage of covered curve points)

```{r echo = FALSE, warning = FALSE, message = FALSE}
# Leave-one curve out approach
# ------------------------------------------------------------------------------
cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
plot_cov_ver(cover.cross.fraction.singlecurve)

@@

covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
covered.curves.percent <- data.frame(t(covered.curves.percent))
colnames(covered.curves.percent) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")

@@

b. (mean percentage of covered curve points)

```{r echo = FALSE, warning = FALSE, message = FALSE}
# Leave-one curve out approach
# ------------------------------------------------------------------------------
cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
plot_cov_ver(cover.cross.fraction.singlecurve)

@@

#### Curves with (missing) peaks

a. (coverage of the entire curve)

```{r echo = FALSE, warning = FALSE, message = FALSE}
data <- example_data(dat = "shock", dir.data)

covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
covered.curves.percent <- data.frame(t(covered.curves.percent))
colnames(covered.curves.percent) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")

b. (mean percentage of covered curve points)

```{r echo = FALSE, warning = FALSE, message = FALSE}
# Leave-one curve out approach
# ------------------------------------------------------------------------------
cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
plot_cov_ver(cover.cross.fraction.singlecurve)

@@

@@

>>>>

### TODO
- Punktweise LoA wasserdicht machen (Parker et al., 2020)
<!-- - Alternativtitel: Comparing different methods for constructing continuous prediction bands -->
<!-- - Exemplarische Paper für die jeweiligen simulierten Datensätze raussuchen -->
- Weitere synthetische Daten siehe Paper Robinson et al. (2021)
- Evtl. simulierte Datensätze um Beispiele aus https://mjskay.github.io/ggdist/articles/lineribbon.html erweitern (insbesondere das erste Beispiel mit x-versetzten Gausskurven)!
- Simulierte Daten durch Funktionen ersetzen? Eigentlich aktuell eher unnötig, da die Zeitreihen ja eigentlich bereits mithilfe von Funktionen modelliert wurden. Zudem wird nicht interpoliert, abgeleitet o. Ä. ... letztlich für die aktuelle Idee des Papers eher ungeeignet (plus man umgeht Probleme wie mögliche Glättungsartefakte, oder die Definition eines functional mean).
- Falls doch FDA, dann evtl. als eigenständiges Skript forken
- Begründung warum Prädiktionsintervalle anstelle von Konfidenzintervallen
<!-- - Umbennung entire curve in "All-or-nothing" (entlehnt von Juul et al.) -->
- .bib file anlegen
- Begriff spatiotemporal überdenken (wie ist das im Kontext FDA definiert)
- Ratkowsky besorgen
- Begründung warum Prädiktionsintervalle und nicht Konfidenzintervalle (theoretisch könnte man beides machen; Prädiktionsintervalle scheinen mir aber eher das zu sein, was man will)

### TODO read (again)
- Parker et al. (2021) https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x
- Robinson, Vanrenterghem, Pataky (2021) Sample size estimation for biomechanical waveforms: Current practice, recommendations and a comparison to discrete power analysis
- Olshen et al. (1989) Gait analysis and the bootstrap
- Curvewise point and interval summaries for tidy data frames of draws from distributions https://mjskay.github.io/ggdist/reference/curve_interval.html
- Horvath, Kokoszka & Reeder (2013) Estimation of the mean of functional time series and a two-sample problem
- Generell die Kooperationspaper von Hörmann und Kokoszka zu FDA Statistik,insbesondere 'Inference for Functional Data with Applications' (2012)
- Ratkowsky 1990 Handbook of nonlinear regression models

@@

source("example_data.R")
# source("fdaDelta.R")
source("pick_subwise_curves.R")
source("draw_clusters.R")
source("floa_rcb.R")
source("floa_point.R")
source("floa_roislien.R")
source("plot_loa.R")
source("get_coverage.R")
source("get_coverage_fraction.R")
source("get_coverage_singlecurve.R")
source("get_coverage_singlecurve_fraction.R")
source("crossval_coverage.R")
source("crossval_coverage_fraction.R")
source("singlecurve_coverage.R")
source("singlecurve_coverage_fraction.R")
source("plot_cov_ver.R")
source("estimate_uncertainty_loa.R")
```

@@

* Randomized Cluster Bootstrap (FLoA~RCB~)
  + One random stride per subject (FLoA~RCB_v2~)
  + Fetch a SINGLE random stride from all strides (FLoA~RCB_v3~)

* Roislien-like intervals (FLoA~Roslien~)
  + Select one random stride from each subject ONCE and bootstrap the resulting sample

Pointwise continuous LoA are constructed using mean +- 1.96 * SD

* Pointwise Gaussian intervals (FLoA~Point~)
  + FLoA~Point~ are calculated using using mean +- 1.96 * SD (standard deviations derived from linear mixed effects models)

Linear mixed effects models (lme4 package (Bates, Maechler, Bolker, & Walker, 2015)) were used because of the advantages over fixed-effects models (e. g. ANOVA as suggested by Bland & Altman, 1999), such as e. g. (i) missing or unbalanced data are less problematic and (ii) inference can be made based on a wider population of patients (Parker et al., 2020). Implementation details are provided in the R code within the supplementary material.

@@

@@

Leave-one (curve) out method to estimate the achieved coverage

Prediction intervals are calculated from size.dataset = number.subjects - 1 and then validated against the left out curve for four different methods:

@@

* FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and bootstrap the resulting sample)

@@

@@

# data <- example_data(dat = "smooth", dir.data)
# floa <- floa_rcb(data, n.boot, ver = "v2")
# plot_loa(data, floa, central.tendency = "mean")
# 
# data <- example_data(dat = "smooth_trend", dir.data)
# floa <- floa_rcb(data, n.boot, ver = "v2")
# plot_loa(data, floa, central.tendency = "mean")
# 
# data <- example_data(dat = "non_gaussian", dir.data)
# floa <- floa_rcb(data, n.boot, ver = "v2")
# plot_loa(data, floa, central.tendency = "mean")
# 
# data <- example_data(dat = "shift", dir.data)
# floa <- floa_rcb(data, n.boot, ver = "v2")
# plot_loa(data, floa, central.tendency = "mean")
```

@@

* 1. FLoA~RCB_v2~ : One random stride per subject

* 2. FLoA~RCB_v3~ : Fetch a SINGLE random stride from all strides

* 3. FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
       strap the resulting sample (of length (n=length(subjects))

* 4. FLoA~Point~ : Pointwise B & A Limits of Agreement (SD derived from linear mixed effects model)

@@

* b: Mean (SD) of curve points within the respective FLoA: The fraction of points along a curve within the limits is calculated and averaged across all curves of a subject.

* c: Uncertainty in LoA (95% percentile intervals)

@@

a. Coverage of the entire curve

@@

covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
covered.curves.percent <- data.frame(t(covered.curves.percent))
colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")

knitr::kable(covered.curves.percent, caption = "Percentages of entirely covered curves")

@@

b. Mean (SD) of covered curve points

```{r echo = FALSE, warning = FALSE, message = FALSE}
# Leave-one curve out approach
# ------------------------------------------------------------------------------
cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)

cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
)
knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")

# # Display as boxplots across methods
# plot_cov_ver(cover.cross.fraction.singlecurve)
```

c. Uncertainty in LoA (95% percentile intervals)

```{r echo = FALSE, warning = FALSE, message = FALSE}
estimate_uncertainty_loa(data, n.boot)
```

@@

a. Coverage of the entire curve

@@

covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
covered.curves.percent <- data.frame(t(covered.curves.percent))
colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")

knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
```

b. Mean (SD) of covered curve points

```{r echo = FALSE, warning = FALSE, message = FALSE}
# Leave-one curve out approach
# ------------------------------------------------------------------------------
cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)

cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
)
knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
```

c. Uncertainty in LoA (95% percentile intervals)

```{r echo = FALSE, warning = FALSE, message = FALSE}
estimate_uncertainty_loa(data, n.boot)
```

@@

covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
covered.curves.percent <- data.frame(t(covered.curves.percent))
colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")

@@

b. Mean (SD) of covered curve points

```{r echo = FALSE, warning = FALSE, message = FALSE}
# Leave-one curve out approach
# ------------------------------------------------------------------------------
cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)

cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
)
knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points")

# # Display as boxplots acroos methods
# plot_cov_ver(cover.cross.fraction.singlecurve)

@@

c. Uncertainty in LoA (95% percentile intervals)

```{r echo = FALSE, warning = FALSE, message = FALSE}
estimate_uncertainty_loa(data, n.boot)
```

@@

c. Uncertainty in LoA (95% percentile intervals)

```{r echo = FALSE, warning = FALSE, message = FALSE}
estimate_uncertainty_loa(data, n.boot)
```

@@

Punktweise FLoA unterrepräsentieren den Messfehler

FLoA Roislien tendieren dazu zu weit auszufallen (den Messfehler zu überrepräsentieren)
